exploratory analysis

### 2023-04-19 03_exploratory.Rmd
Create 03_exploratory.Rmd

Required Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sf)
library(rerddap)
library(terra)
library(marmap)
library(rnaturalearth)
library(mapproj)
library(raster)
library("rerddapXtracto")
library(PAMpal)
library(PAMmisc)
library(maptiles)
library(tmap)
library(spData)
library(tmaptools)
library(grid)
library(ggspatial)
```

Set Parameters
```{r}
theme_set(theme_bw())
```

### Exploratory analysis

#### Not important: tracks as sf and basemaps

```{r}
library(tmap); library(maptiles); library(sf)
tmap_mode("plot")
trackssf <- st_as_sf(tracks, coords = c("long", "lat"), crs=4326)
#tracks08sf <- st_as_sf(track_BB_OL_08, coords = c("long", "lat"), crs=4326)  # outdated
#tracks10sf <- st_as_sf(track_BB_OL_10, coords = c("long", "lat"), crs=4326)  # outdated
trk_TOL_08 <- st_as_sf(tracks_SS_08, coords = c("long", "lat.x"), crs=4326)
trk_TOL_10 <- st_as_sf(tracks_SS_10, coords = c("long", "lat.x"), crs=4326)
trackBase <- get_tiles(whales_8_10sf, provider = "Esri.WorldShadedRelief")
allTrackbase <- get_tiles(trackssf, provider = "Stamen.TonerLite", zoom = 5)
```

### Sea Surface Temp with rerddap, Drift 8 only
```{r}
# Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Daily
SSTInfo <- info('jplMURSST41') 
murSST <- griddap(SSTInfo, latitude = c(35, 39), longitude = c(-122, -130), 
                  time = c("2018-08-16", "2018-10-01"), 
                  fields = 'analysed_sst') # drift8 dates only
SST <- murSST$data

### SST from NOAA CoastWatch : https://coastwatch.pfeg.noaa.gov/projects/r/xyt-matchup.html
```{r}
xcoord <- tracks_08$long
ycoord <- tracks_08$lat
tcoord <- tracks_08$dateTime
xlen <- 0.2
ylen <- 0.2
parameter <- "analysed_sst"
dataset <- 'jplMURSST41'
#dataInfo <- rerddap::info(dataset, url= "https://coastwatch.pfeg.noaa.gov/erddap/")

#swchl <- rxtracto(dataInfo, 
#                  parameter=parameter, 
#                  xcoord=xcoord, ycoord=ycoord, 
#                  tcoord=tcoord, xlen=xlen, ylen=ylen)

#plotTrack(swchl, xcoord, ycoord, tcoord, plotColor = 'thermal')
```

### SST from PAMPal
```{r}
# Downloads data from servers
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'jplMURSST41', fileName = "analysedSST") %>%

# from file 
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/analysedSST.nc') %>%
  rename(., sst_mean = analysed_sst_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)   # rename new column to "sst_mean" and remove unnecessary columns
tracks_SS_08
```

### Distance to continental slope

```{r}
# Increase cell size by 30x to be visible
bathy <- rast("data/gebco_bathy.tif")
agg <- terra::aggregate(bathy, 20)
plot(agg)

# Rasterize track with agg as the template
#track_rast <- rasterize(tracks08sf, agg, field = "BB_20-24000")
#TOL_rast <- rasterize(trk_TOL_08, agg, field = "TOL_125")
sf08 <- st_as_sf(tracks_SS_08, coords = c("long", "lat.x"), crs=4326, remove=F)
sf10 <- st_as_sf(tracks_SS_10, coords = c("long", "lat.x"), crs=4326, remove=F)
trkraster08 <- rasterize(vect(sf08), agg)
trkraster10 <- rasterize(vect(sf10), agg)

@@

ggplot() + 
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y, 
                            alpha = 0.8,
                            fill = `gebco_bathy`))+
  geom_sf(data = cont, col="black") + 
  geom_sf(data = coast, col = "red") +
  geom_sf(data = cont_slope, col = "purple") +
  geom_sf(data = whales_8_10sf, mapping = aes()) +
  scale_color_gradient(low="green", high="red") 
```

### Extract slope_dist and bind to drift tracks
```{r}
# Drift 8
slopedist <- terra::distance(trkraster08, vect(cont_slope))     # measure distance from track to slope
extdist <- terra::extract(slopedist, vect(sf08))       # extract this distance into a df
tracks_SS_08 <- cbind(sf08,extdist) %>%                     # bind to main df
  dplyr::select(-ID) %>%
  rename("dist_slope" = gebco_bathy)

#Drift 10
slopedist <- terra::distance(trkraster10, vect(cont_slope))     # measure distance from track to slope
extdist <- terra::extract(slopedist, vect(sf10))       # extract this distance into a df
tracks_SS_10 <- cbind(sf10,extdist) %>%                     # bind to main df
  dplyr::select(-ID) %>%
  rename("dist_slope" = gebco_bathy)
```

#### Plot distance to continental slope

```{r}
# By color
#ggplot() +
#  geom_sf(data = pts08, 
#          mapping = aes(col = `Distance to Continental Slope`)) +
#  geom_sf(data = pts10, 
#          mapping = aes(col = `Distance to Continental Slope`)) +
#  geom_sf(data = cont_slope)

# Plot distance by OL/TOL 
#ggplot(data = pts08, aes(x = `Distance to Continental Slope`, y = `OL_125`)) + geom_point()
#ggplot(data = TOL08, aes(x = `Distance to Continental Slope`, y = `TOL_125`)) + geom_point() + geom_smooth()
   # what are the units for distance? Meters

ggplot() +
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y,
                            fill = `gebco_bathy`)) +
  geom_sf(data = tracks_SS_08, 
          mapping = aes(col = dist_slope)) +
  geom_sf(data = cont_slope) + scale_fill_continuous(high = "lightgreen", low = "blue")

@@

```{r drift-8-histogram}
```

@@

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz")
# TOL linear model plots
#plot(lm(formula = `Distance to Continental Slope` ~ bestNumber, data = TOL_whale_08))

#ggplot(TOL_whale_08) +
#  geom_point(aes(x = TOL_125, y = nClicks), color = "blue") #+
#  geom_point(aes(x = TOL_63, y =  nClicks), color = "hotpink") +
#  geom_point(aes(x = TOL_5000, y =  nClicks), color = "limegreen") +
#  geom_point(aes(x = TOL_2000, y =  nClicks), color = "yellow")

### Does hour of day affect sound levels?

```{r}
# CCES_08
#hr_08 <- track_BB_OL_08 %>%
#  mutate(Hr = hour(track_BB_OL_08$dateTime)) %>%
#  group_by(Hr) %>%
#  summarize(meanBB = mean(`BB_20-24000`)) %>%
#  rename(`Buoy 08 Broadband` = `meanBB`)
#ggplot(data = hr_08, mapping = aes(x = Hr, y = `Buoy 08 Broadband`)) + 
#  geom_point() + geom_smooth()

#CCES_10
#hr_10 <- track_BB_OL_10 %>%
#  mutate(Hr = hour(track_BB_OL_10$dateTime)) %>%
#  group_by(Hr) %>%
#  summarize(meanBB = mean(`BB_20-24000`)) %>%
#  rename(`Buoy 10 Broadband` = `meanBB`)
#ggplot(data = hr_10, mapping = aes(x = Hr, y = `Buoy 10 Broadband`)) + 
#  geom_point() + geom_smooth()

# Join and graph
#hr_join <- left_join(hr_08, hr_10, by = "Hr")
#hr_piv <- hr_join %>%
#  pivot_longer(col=c(`Buoy 08 Broadband`,`Buoy 10 Broadband`), names_to = "buoys", values_to = "BB")

#ggplot(data = hr_piv, mapping = aes(Hr, BB, col=buoys)) +
#  geom_point() + geom_smooth()
```

### Are sound levels affected by species composition or number of clicks?

```{r}
# Plot
library(tmap); library(maptiles)
tmap_mode("plot")
oceanBase <- get_tiles(tracks_SS_08, provider="Esri.NatGeoWorldMap", crop = TRUE)
tm_shape(oceanBase) + tm_rgb() +
  tm_shape(tracks_SS_08) +  tm_symbols(col="species", size = 1) +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"))

# CCES 08
#BB_whales08 <- st_join(pts08, whales_sf) %>%
#  filter(!is.na(`species`))

#summary(glm(`BB_20.24000` ~ species + `Distance to Continental Slope` + nClicks, data = BB_whales08))

# CCES 10
#BB_whales10 <- st_join(pts10, whales_sf) %>%
#  filter(!is.na(`species`))

#summary(glm(`BB_20.24000` ~ species + `Distance to Continental Slope` + nClicks, data = BB_whales10))

#Raster base map
tmap_mode("plot")
oceanBase <- get_tiles(tracks_SS_10, provider="Esri.NatGeoWorldMap")
tm_shape(bathy) + tm_raster() +
  tm_shape(tracks_SS_10) + 
  tm_symbols(style = "jenks") +
  tm_layout(legend.width = 0.5)

#### Bathymetry

```{r}
bgo <- getNOAA.bathy(lon1 = -121, lon2 = -131,
                     lat1 = 34, lat2= 40, resolution = 1)  # plot bounding box around drift 8 and 10
# As raster
bgo_rast <- marmap::as.raster(bgo)

# Convert sf to dataframe for this operation
tracks_SS_08 <- st_set_geometry(tracks_SS_08, value = NULL)

# Get depth of whale detections in drift 8
depth08 <- get.depth(bgo, tracks_SS_08[,3:2], locator = FALSE)

# Get depth of tracks
trkdepth08 <- get.depth(bgo, tracks_SS_08[,3:2], locator = FALSE)
trkjoin <- left_join(tracks_SS_08, trkdepth08, by = c("long"= "lon")) %>%
  mutate(depth = depth*-1)         # remove the negative sign

#Plot
ggplot() + geom_point(data = tracks_SS_08, aes(x = long, y = lat.x, size = 2)) +
  scale_color_gradient(low = "red", high = "beige") +
  ggtitle("Depth of Drift 8 whale detections")
```
#### Bathymetry - PAMpal
```{r}
# Downloads from server
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'srtm15plus', fileName='depth.nc')

# Pulls from downloaded file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/depth.nc') %>%
  rename(., depth = z_mean) %>%        # rename new column to "depth"
  mutate(depth = depth*-1) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)
tracks_SS_08
```

```{r}
# Slope
bgo_slope <- terrain(bgo_rast, v = "slope")
plot(bgo_slope, axes = FALSE)

# Extract slope values at points along drift 8
slope <- raster::extract(bgo_slope, tracks_SS_08[,3:2])
tracks_SS_08 <- cbind(tracks_SS_08, slope)

#ggplot() +
#  geom_raster(data = bgo_slope, mapping = aes(x=x, y=y, fill=slope)) + 
#  geom_point(data = depth08, aes(x = lon, y = lat, color = depth, size = 2))

# Roughness (different from rugosity?)
plot(terrain(bgo_rast, opt = "roughness"))    # what are the units on this?
```

#### Distance from coast - PAMPal

```{r}
dist2shore <- erddapToEdinfo(dataset='dist2coast_1deg',
                             baseurl='https://pae-paha.pacioos.hawaii.edu/erddap/')
# Downloads nc file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc=dist2shore, fileName='Dist2Shore.nc')

tracks_SS_08 <- matchEnvData(tracks_SS_08, nc='data/Dist2Shore.nc') %>%
  rename(., dist_shore = dist_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)
tracks_SS_08
```

#### Chlorophyll-a: NOAA Coastwatch

```{r}
# Code chunk taken from NOAA CoastWatch : https://coastwatch.pfeg.noaa.gov/projects/r/xyt-matchup.html
xcoord <- tracks_SS_08$long
ycoord <- tracks_SS_08$lat
tcoord <- tracks_SS_08$dateTime
xlen <- 0.2
ylen <- 0.2
parameter <- "chlorophyll"
dataset <- 'erdMH1chlamday'
dataInfo <- rerddap::info(dataset, url= "https://coastwatch.pfeg.noaa.gov/erddap/")

swchl <- rxtracto(dataInfo, 
                  parameter=parameter, 
                  xcoord=xcoord, ycoord=ycoord,
                  tcoord=tcoord, xlen=xlen, ylen=ylen)

plotTrack(swchl, xcoord, ycoord, tcoord, plotColor = 'algae')

# Pulls from downloaded file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/chlorophyll.nc')
tracks_SS_08
```

#### Sea surface height anomalies - NOAA Coastwatch
```{r}
# Code chunk taken from NOAA CoastWatch : https://coastwatch.pfeg.noaa.gov/projects/r/xyt-matchup.html
xcoord <- tracks_SS_08$long
ycoord <- tracks_SS_08$lat
tcoord <- tracks_SS_08$dateTime
xlen <- 0.2
ylen <- 0.2
parameter <- 'sla'
dataset <- 'nesdisSSH1day'
dataInfo <- rerddap::info(dataset, url= "https://coastwatch.pfeg.noaa.gov/erddap/")

swssh <- rxtracto(dataInfo, 
                  parameter=parameter, 
                  xcoord=xcoord, ycoord=ycoord, 
                  tcoord=tcoord, xlen=xlen, ylen=ylen)

plotTrack(swssh, xcoord, ycoord, tcoord, plotColor = 'deep')
```

# Pulled from downloaded nc file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/sshanomalies.nc') %>%
  rename(., ssh_mean = sla_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)
tracks_SS_08
```

### Wind Stress Curl from PAMPal *** NEEDS FIXING ****
```{r}
# Downloads from server
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdQSstress1day_LonPM180', #fileName='erdQSstress1day_LonPM180.nc', var = 'curl', depth = 0)

# Pulls from downloaded file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdQSstress1day_LonPM180.nc')
#tracks_SS_08

#windcurl <- erddapToEdinfo('erdQSstress1day_LonPM180', baseurl = 'https://coastwatch.pfeg.noaa.gov/erddap/')

#### HyCom Data - MLDepth, MLTemp, TTemp, TDepth, Salinity
```{r}
# These refer to nc files that have already been downloaded from the server

#hycomData0 <- matchEnvData(as.data.frame(tracks_SS_08[1:250,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = #"hycomm0.nc")

#hycomData1 <- matchEnvData(as.data.frame(tracks_SS_08[251:500,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = #"hycomm1.nc")

#hycomData2 <- matchEnvData(as.data.frame(tracks_SS_08[501:750,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = #"hycomm2.nc")

#hycomData3 <- matchEnvData(as.data.frame(tracks_SS_08[751:1200,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName #= "hycomm3.nc")

#hycomData4 <- matchEnvData(as.data.frame(tracks_SS_08[1201:1500,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName #= "hycomm4.nc")

#hycomData5 <- matchEnvData(as.data.frame(tracks_SS_08[1501:1800,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName #= "hycomm5.nc")

#hycomData6 <- matchEnvData(as.data.frame(tracks_SS_08[1801:2145,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName #= "hycomm6.nc")

# Run the 7 nc lists that make up tracks_SS_08
hycomData0 <- matchEnvData(as.data.frame(tracks_SS_08[1:250,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm0.nc")
hycomData1 <- matchEnvData(as.data.frame(tracks_SS_08[251:500,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm1.nc")
hycomData2 <- matchEnvData(as.data.frame(tracks_SS_08[501:750,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm2.nc")
hycomData3 <- matchEnvData(as.data.frame(tracks_SS_08[751:1200,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm3.nc")
hycomData4 <- matchEnvData(as.data.frame(tracks_SS_08[1201:1500,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm4.nc")
hycomData5 <- matchEnvData(as.data.frame(tracks_SS_08[1501:1800,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm5.nc")
hycomData6 <- matchEnvData(as.data.frame(tracks_SS_08[1801:2140,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm6.nc")
```

Use function to estimate thermocline variables using the temperature at depth data
```{r}
calcThermoVars <- function(depth, temp, mldMode='SST8', thermMode='variso', plot=FALSE) {
    # depthInterp <- approxfun(x=temp, y=depth)
    switch(match.arg(mldMode),
           'SST8' = {
               mldTemp <- temp[1] - 0.8
               # first few temps are likely to have repeat values that trigger
               # warnings for approxfun, so cut off to only values after mldTemp
               useMin <- min(which(temp <= mldTemp)) - 1
               useIx <- useMin:length(temp)
               depthInterp <- approxfun(x=temp[useIx], y=depth[useIx])
               mld <- depthInterp(mldTemp)
           }
    )
    switch(match.arg(thermMode),
           'variso' = {
               t400 <- approx(x=depth, y=temp, xout=400)$y
               # therm end is half way to t400 temp
               tEnd <- mldTemp + (t400 - mldTemp)/2
               # therm temp is midpt from mld to end
               tt <- (mldTemp + tEnd)/2
               # tt <- mldTemp - .25 * (mldTemp - t400)
               ttDepth <- depthInterp(tt)
               tEndDepth <- depthInterp(tEnd)
           }
    )
    if(plot) {
        to300 <- depth <= 300
        plot(x=temp[to300], y=-depth[to300],
             main = paste0('MixLayer: ', round(mld,0), 'm ', round(mldTemp, 1), 'C',
                           '\nTherm: ', round(ttDepth, 0), 'm ', round(tt, 1), 'C'),
             xlab='Temp (C)', ylab='Depth (m)')

```{r}
#calcThermoVars(hycomData[[1]]$matchDepth, hycomData[[1]]$water_temp, plot=TRUE)

thermData <- rbind(thermData0, thermData1, thermData2, thermData3, thermData4, thermData5, thermData6)

tracks_SS_08 <- cbind(tracks_SS_08, thermData)

# New tracks_SS_08 variable contains Hycom variables now
View(tracks_SS_08)
```

>>>>

Required Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sf)
library(rerddap)
library(terra)
library(marmap)
library(rnaturalearth)
library(mapproj)
library(raster)
#library("rerddapXtracto")
library(PAMpal)
library(PAMmisc)
library(maptiles)
library(tmap)
library(spData)
library(tmaptools)
library(grid)
library(ggspatial)
```

Set Parameters
```{r}
theme_set(theme_bw())
tmap_mode("plot")
```

Create sf objects and basemaps for maps
```{r}
whales_8_10sf <- BWsf %>%                            
dplyr::filter(station == "8" | station == "10")

trackssf <- st_as_sf(tracks, coords = c("Longitude", "Latitude"), crs=4326)
trk_TOL_08 <- st_as_sf(tracks_SS_08, coords = c("Longitude", "Latitude"), crs=4326)
trk_TOL_10 <- st_as_sf(tracks_SS_10, coords = c("Longitude", "Latitude"), crs=4326)
trackBase <- get_tiles(whales_8_10sf, provider = "Esri.WorldShadedRelief")
allTrackbase <- get_tiles(trackssf, provider = "Stamen.TonerLite", zoom = 5)
```
### Exploratory analysis

SST from PAMPal
```{r}
# Downloads data from servers, uncomment and only run once
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'jplMURSST41', fileName = "analysedSST")

# from file 
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/analysedSST.nc') %>%
  rename(., sst_mean = analysed_sst_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)   # rename new column to "sst_mean" and remove unnecessary columns
```

Distance to continental slope
```{r}
# Increase cell size by 30x to be visible
bathy <- rast("data/gebco_bathy.tif")
agg <- terra::aggregate(bathy, 5)
plot(agg)

# Rasterize track with agg as the template

#TOL_rast <- rasterize(trk_TOL_08, agg, field = "TOL_125")
sf08 <- st_as_sf(tracks_SS_08, coords = c("Longitude", "Latitude"), crs=4326, remove=F)
sf10 <- st_as_sf(tracks_SS_10, coords = c("Longitude", "Latitude"), crs=4326, remove=F)
trkraster08 <- rasterize(vect(sf08), agg)
trkraster10 <- rasterize(vect(sf10), agg)

@@

#ggplot() + 
#  geom_raster(data = bathy_df, 
#              mapping = aes(x=x, y=y, 
#                            alpha = 0.8,
#                            fill = `gebco_bathy`))+
#  geom_sf(data = cont, col="black") + 
#  geom_sf(data = coast, col = "red") +
#  geom_sf(data = cont_slope, col = "purple") +
#  geom_sf(data = whales_8_10sf, mapping = aes()) +
#  scale_color_gradient(low="green", high="red") 
```

### Extract slope_dist and bind to drift tracks
```{r}
# Drift 8
slopedist <- terra::distance(trkraster08, vect(cont_slope))     # measure distance from track to slope
extdist <- terra::extract(slopedist, vect(sf08))       # extract this distance into a df
tracks_SS_08 <- cbind(sf08,extdist) %>%                     # bind to main df
  dplyr::select(-ID) %>%
  rename("dist2slope" = gebco_bathy) %>%
  st_set_geometry(NULL)                         # Drop geometry to make df for future cbinds

#Drift 10
slopedist <- terra::distance(trkraster10, vect(cont_slope))     # measure distance from track to slope
extdist <- terra::extract(slopedist, vect(sf10))       # extract this distance into a df
tracks_SS_10 <- cbind(sf10,extdist) %>%                     # bind to main df
  dplyr::select(-ID) %>%
  #rename("dist2slope" = gebco_bathy) %>%
  st_set_geometry(NULL)                        # Drop geometry to make df for future cbinds
```

#### Bathymetry - PAMpal
```{r}
# Downloads from server
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'srtm15plus', fileName='depth.nc')

# Pulls from downloaded file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/depth.nc') %>%
  rename(., depth = z_mean) %>%        # rename new column to "depth"
  mutate(depth = depth*-1) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)
```

```{r}
# Slope (Hard-coded - can fix this)
bgo <- getNOAA.bathy(lon1 = -121, lon2 = -131,
                     lat1 = 34, lat2= 40, resolution = 1)  # plot bounding box around drift 8 and 10
# As raster
bgo_rast <- marmap::as.raster(bgo)
bgo_slope <- terrain(bgo_rast, v = "slope")

# Extract slope values at points along drift 8
slope <- raster::extract(bgo_slope, tracks_SS_08[,4:3])
tracks_SS_08 <- cbind(tracks_SS_08, slope)
```

#### Distance from coast - PAMPal
Need to have variable choose itself!! Also, how do I get it to not download each time?
```{r}
#dist2shore <- erddapToEdinfo(dataset='dist2coast_1deg',
#                             baseurl='https://pae-paha.pacioos.hawaii.edu/erddap/',
#                             chooseVars = FALSE)

tracks_SS_08 <- matchEnvData(tracks_SS_08, nc='data/Dist2Shore.nc') %>%
  rename(., dist2shore = dist_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

#May not need this part
# Downloads from server
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'dist2coast_1deg', fileName='dist2coast.nc', var = 'dist')

# Pulls from downloaded file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/dist2coast.nc')

# Downloads nc file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc=dist2shore, fileName='Dist2Shore.nc')  
```

# Pulls from downloaded file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/chlorophyll.nc')
```

# Pulled from downloaded nc file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data/sshanomalies.nc') %>%
  rename(., ssh_mean = sla_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)
```

### Wind Stress Curl from PAMPal *** NEEDS FIXING ****
```{r}
# Downloads from server
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdQMstress1day_LonPM180', fileName='erdQMstress1day_LonPM180.nc', var = 'curl')

# Pulls from downloaded file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdQMstress1day_LonPM180.nc')
#tracks_SS_08

#windcurl <- erddapToEdinfo('erdQMstress1day', baseurl = 'https://coastwatch.pfeg.noaa.gov/erddap/')

#### HyCOM Data - MLDepth, MLTemp, TTemp, TDepth, Salinity
Doesn't bin itself into separate dataframe based on size!
```{r}
# These refer to nc files that have already been downloaded from the server.
# Uncomment and only run once! Rows are hard-coded.

#hycomData0 <- matchEnvData(as.data.frame(tracks_SS_08[1:250,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm0.nc")

#hycomData1 <- matchEnvData(as.data.frame(tracks_SS_08[251:500,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm1.nc")

#hycomData2 <- matchEnvData(as.data.frame(tracks_SS_08[501:750,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm2.nc")

#hycomData3 <- matchEnvData(as.data.frame(tracks_SS_08[751:1200,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm3.nc")

#hycomData4 <- matchEnvData(as.data.frame(tracks_SS_08[1201:1500,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm4.nc")

#hycomData5 <- matchEnvData(as.data.frame(tracks_SS_08[1501:1800,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm5.nc")

#hycomData6 <- matchEnvData(as.data.frame(tracks_SS_08[1801:2145,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm6.nc")

#hycomData7 <- matchEnvData(as.data.frame(tracks_SS_08[2146:2469,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm7.nc")

#hycomData8 <- matchEnvData(as.data.frame(tracks_SS_08[2470:2769,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "hycomm8.nc")

# Run the 9 nc lists that make up tracks_SS_08
hycomData0 <- matchEnvData(as.data.frame(tracks_SS_08[1:250,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm0.nc")
hycomData1 <- matchEnvData(as.data.frame(tracks_SS_08[251:500,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm1.nc")
hycomData2 <- matchEnvData(as.data.frame(tracks_SS_08[501:750,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm2.nc")
hycomData3 <- matchEnvData(as.data.frame(tracks_SS_08[751:1200,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm3.nc")
hycomData4 <- matchEnvData(as.data.frame(tracks_SS_08[1201:1500,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm4.nc")
hycomData5 <- matchEnvData(as.data.frame(tracks_SS_08[1501:1800,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm5.nc")
hycomData6 <- matchEnvData(as.data.frame(tracks_SS_08[1801:2145,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm6.nc")
hycomData7 <- matchEnvData(as.data.frame(tracks_SS_08[2146:2469,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm7.nc")
hycomData8 <- matchEnvData(as.data.frame(tracks_SS_08[2470:2769,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "hycomm8.nc")
```

Create function to estimate thermocline variables using the temperature at depth data
```{r}
calcThermoVars <- function(depth, temp, mldMode='SST8', thermMode='variso', plot=FALSE) {
    # depthInterp <- approxfun(x=temp, y=depth)
    switch(match.arg(mldMode),
           'SST8' = {
               mldTemp <- temp[1] - 0.8
               # first few temps are likely to have repeat values that trigger
               # warnings for approxfun, so cut off to only values after mldTemp
               useMin <- min(which(temp <= mldTemp)) - 1
               useIx <- useMin:length(temp)
               depthInterp <- approxfun(x=temp[useIx], y=depth[useIx])
               mld <- depthInterp(mldTemp)
           }
    )
    switch(match.arg(thermMode),
           'variso' = {
               t400 <- approx(x=depth, y=temp, xout=400)$y
               # therm end is half way to t400 temp
               tEnd <- mldTemp + (t400 - mldTemp)/2
               # therm temp is midpt from mld to end
               tt <- (mldTemp + tEnd)/2
               # tt <- mldTemp - .25 * (mldTemp - t400)
               ttDepth <- depthInterp(tt)
               tEndDepth <- depthInterp(tEnd)
           }
    )
    if(plot) {
        to300 <- depth <= 300
        plot(x=temp[to300], y=-depth[to300],
             main = paste0('MixLayer: ', round(mld,0), 'm ', round(mldTemp, 1), 'C',
                           '\nTherm: ', round(ttDepth, 0), 'm ', round(tt, 1), 'C'),
             xlab='Temp (C)', ylab='Depth (m)')

Use above function on the smaller hycom datasets (0-6)
```{r}
#calcThermoVars(hycomData[[1]]$matchDepth, hycomData[[1]]$water_temp, plot=TRUE)

thermData7 <- bind_rows(lapply(hycomData7, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData8 <- bind_rows(lapply(hycomData8, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

# Bind all thermocline datasets together
thermData <- rbind(thermData0, thermData1, thermData2, thermData3, thermData4, 
                   thermData5, thermData6, thermData7, thermData8)

# Bind main thermocline dataset to main track/soundscape dataframe
tracks_SS_08 <- cbind(tracks_SS_08, thermData)

# New tracks_SS_08 variable contains Hycom variables now
##### FINAL tracks_ss_08 dataframe ######
View(tracks_SS_08)
```

Get avg daily environmental variables
```{r}
means_08 <- tracks_SS_08 %>%
  group_by(as.Date(UTC))%>%
  summarize(sst_mean = mean(sst_mean),
            dist_slope = mean(dist2slope),
            depth = mean(depth),
            slope = mean(slope),
            dist_shore = mean(dist2shore),
            chlorophyll_mean = mean(chlorophyll_mean),
            ssh_mean = mean(ssh_mean),
            mldDepth = mean(mldDepth),
            mldTemp = mean(mldTemp),
            ttDepth = mean(ttDepth),
            ttTemp = mean(ttTemp),
            temp400 = mean(temp400),
            sal400 = mean(sal400),
            BWpresence = sum(BWpresence))
```

### Exploratory Analysis ######

@@

@@
